---
title: "Intro to R and RStudio"
subtitle: "A Quick Start for Humanities Scholars"
author: "Dr Giulia Grisot"
date: "May 28, 2025"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Welcome!

## Objectives for This Intro

- Understand what **R** and **RStudio** are
- Get familiar with the RStudio interface
- Try some basic R code
- Prepare to work with R Markdown

---

# What is R? What is RStudio?

- **R** is a programming language for data analysis and visualization.
- **RStudio** is an *IDE* — Integrated Development Environment — for using R more easily.

> Think of R as the engine, and RStudio as the dashboard.

---

# RStudio: The Interface

When you open RStudio, you'll see **four panes**:

| Pane | Purpose | Example |
|------|---------|---------|
| Console | Run code, see results | `2 + 2` |
| Script Editor | Write reusable code | Save as `.R` or `.Rmd` |
| Environment / History | View data & past code | `x <- c(1,2,3)` |
| Files / Plots / Help | Navigate, install, get help | `plot(1:10)` |

---

# Try This: Console

Type the following in the **Console**:

```{r}
2 + 2
sqrt(25)
mean(c(10, 20, 30))
```

---

# Try This: Script Editor

Open a new R Script (File → New File → R Script) and try:

```{r}
# This is a comment
x <- c(1, 2, 3, 4, 5)
mean(x)
```

- Click **Run** to execute the line
- Notice how the result appears in the Console
- Save your script as `my_first_script.R`

---

# Install and Load a Package

Try installing a package (if you have internet):

```{r eval=FALSE}
install.packages("tidyverse")
library(tidyverse)
```

- Install: one-time action per machine
- Load (`library()`): every time you want to use it

---

# Bonus: View a Plot

Try making a quick plot:

```{r}
plot(1:10)
```

It should appear in the **Plots** pane (bottom-right).

---

# Save Your Work

1. Save your `.R` script to a folder on your computer.
2. Use File → Save or `Cmd/Ctrl + S`
3. You’ll be working inside a **Project** soon — easier to manage!

---

# Checkpoint: What Have You Learned?

- Difference between Console and Script?
- What does `x <- 2 + 2` do?
- What’s the benefit of writing in the Script Editor?

---

# Next Up: R Markdown

We’ll now move into **R Markdown**, where you can combine narrative, code, and output in a single document — perfect for research and teaching.

Let’s get started!


